---
title: "Linear Marketing Mix Model Benchmark"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

This notebook contains the linear-regression benchmark developed before the
final nonlinear Marketing Mix Model.

The original proprietary dataset is not included. Company, brand and campaign
names have been anonymized.

## Packages and data

```{r packages}
library(readxl)
library(dplyr)
library(ggplot2)
library(prophet)
library(MASS)
library(car)
library(lmtest)
library(sandwich)
library(caret)
```

```{r data}
data_path <- "data/private/marketing_data.xlsx"

if (!file.exists(data_path)) {
  stop("Private dataset not found. The proprietary dataset used in the thesis is not distributed with this repository.")
}

df <- read_excel(data_path, sheet = "model_data")
```

## Seasonality

```{r seasonality}
prophet_data <- df %>%
  select(Week_Ending, Vol_Sales) %>%
  rename(ds = Week_Ending, y = Vol_Sales)

m <- prophet(
  prophet_data,
  yearly.seasonality = TRUE,
  weekly.seasonality = TRUE,
  daily.seasonality = FALSE,
  seasonality.mode = "multiplicative",
  growth = "linear"
)

forecast <- predict(m, prophet_data)
seasonal_component <- forecast$yearly + forecast$weekly

df$seasonal_index_prophet <- exp(seasonal_component) / mean(exp(seasonal_component))
```

## Pre-processing

```{r preprocessing}
df$search <- df$CLICK_Search_Channel_A + df$CLICK_Search_Channel_B

correlation_vars <- c(
  "Vol_Sales", "GRP_TV_target", "GRP_TV_related_1", "GRP_TV_related_2",
  "imp_display", "competition", "search", "imp_video_social",
  "imp_meta_youtube", "seasonal_index_prophet"
)

cor_matrix <- cor(df[, correlation_vars], use = "complete.obs")
print(cor_matrix)

Q1 <- quantile(df$Vol_Sales, 0.25)
Q3 <- quantile(df$Vol_Sales, 0.75)
IQR_sales <- Q3 - Q1

lower_bound <- Q1 - 1.5 * IQR_sales
upper_bound <- Q3 + 1.5 * IQR_sales

outliers <- df$Vol_Sales[df$Vol_Sales < lower_bound | df$Vol_Sales > upper_bound]
print(paste("Number of outliers:", length(outliers)))
```

## Adstock preparation

```{r adstock}
adstock <- function(x, decay_rate) {
  result <- numeric(length(x))
  result[1] <- x[1]

  for (i in 2:length(x)) {
    result[i] <- x[i] + decay_rate * result[i - 1]
  }

  return(result)
}

tv_decay <- 0.8
digital_decay <- 0.5

df$GRP_TV_target_adstock <- adstock(df$GRP_TV_target, tv_decay)
df$GRP_TV_related_1_adstock <- adstock(df$GRP_TV_related_1, tv_decay)
df$GRP_TV_related_2_adstock <- adstock(df$GRP_TV_related_2, tv_decay)

df$imp_display_adstock <- adstock(df$imp_display, digital_decay)
df$imp_video_social_adstock <- adstock(df$imp_video_social, digital_decay)
df$imp_meta_youtube_adstock <- adstock(df$imp_meta_youtube, digital_decay)
df$search_adstock <- adstock(df$search, digital_decay)
```

## Initial linear model

```{r model-1}
model1 <- lm(
  Vol_Sales ~
    Base_price +
    NEW_DATA_NMR +
    Promo_Intensity +
    seasonal_index_prophet +
    GRP_TV_target_adstock +
    GRP_TV_related_1_adstock +
    GRP_TV_related_2_adstock +
    competition +
    imp_video_social_adstock +
    imp_meta_youtube_adstock +
    imp_display_adstock +
    search_adstock,
  data = df
)

summary(model1)
vif(model1)
plot(model1)

resettest(model1, power = 2, type = "fitted", data = df)
ncvTest(model1)
bptest(model1)
```

## Log-transformed model with selected interactions

```{r model-2}
model2 <- lm(
  log(Vol_Sales) ~
    Base_price +
    NEW_DATA_NMR +
    GRP_TV_target_adstock +
    GRP_TV_related_1_adstock +
    GRP_TV_related_2_adstock +
    seasonal_index_prophet * competition +
    imp_video_social_adstock +
    imp_meta_youtube_adstock +
    Promo_Intensity * imp_display_adstock +
    search_adstock,
  data = df
)

summary(model2)
vif(model2)
plot(model2)

resettest(model2, power = 2, type = "fitted", data = df)
ncvTest(model2)
bptest(model2)

robust_se <- coeftest(model2, vcov = vcovHC(model2, type = "HC3"))
print(robust_se)
```

## Standardized media benchmark

```{r standardized-media}
standardize_no_zeros <- function(x) {
  non_zero <- x[x != 0]

  if (length(non_zero) > 0) {
    return((x - mean(non_zero)) / sd(non_zero))
  } else {
    return(x)
  }
}

media_vars <- c(
  "GRP_TV_target_adstock",
  "GRP_TV_related_1_adstock",
  "GRP_TV_related_2_adstock",
  "imp_display_adstock",
  "imp_video_social_adstock",
  "imp_meta_youtube_adstock",
  "search_adstock",
  "competition"
)

for (var in media_vars) {
  new_var_name <- paste0(var, "_std")
  df[[new_var_name]] <- standardize_no_zeros(df[[var]])
}
```

```{r final-linear-benchmark}
model_std2 <- lm(
  log(Vol_Sales) ~
    NEW_DATA_NMR +
    I(Base_price^2) +
    GRP_TV_target_adstock_std +
    GRP_TV_related_1_adstock_std +
    GRP_TV_related_2_adstock_std +
    Promo_Intensity * imp_display_adstock_std +
    imp_video_social_adstock_std +
    imp_meta_youtube_adstock_std +
    search_adstock_std +
    competition_std * seasonal_index_prophet,
  data = df
)

summary(model_std2)
plot(model_std2)

resettest(model_std2, power = 2, type = "fitted", data = df)
ncvTest(model_std2)
bptest(model_std2)

robust_se <- coeftest(model_std2, vcov = vcovHC(model_std2, type = "HC3"))
print(robust_se)
```

## Time-series cross-validation

```{r validation}
build_model <- function(data) {
  lm(
    log(Vol_Sales) ~
      log(NEW_DATA_NMR) +
      Base_price +
      GRP_TV_target_adstock_std +
      GRP_TV_related_1_adstock_std +
      GRP_TV_related_2_adstock_std +
      Promo_Intensity * imp_display_adstock_std +
      imp_video_social_adstock_std +
      imp_meta_youtube_adstock_std +
      search_adstock_std +
      seasonal_index_prophet * competition_std,
    data = data
  )
}

df$date <- as.Date(df$Week_Ending)
df <- df[order(df$date), ]

initial_window <- 52
horizon <- 13

indices <- createTimeSlices(
  1:nrow(df),
  initialWindow = initial_window,
  horizon = horizon,
  fixedWindow = TRUE
)

rmse <- function(actual, predicted) {
  sqrt(mean((actual - predicted)^2))
}

cv_results <- list()

for (i in 1:length(indices$train)) {
  train_data <- df[indices$train[[i]], ]
  test_data <- df[indices$test[[i]], ]

  model <- build_model(train_data)
  predictions <- exp(predict(model, newdata = test_data))
  actual <- test_data$Vol_Sales

  cv_results[[i]] <- list(
    rmse = rmse(actual, predictions),
    mape = mean(abs((actual - predictions) / actual)) * 100
  )
}

mean_rmse <- mean(sapply(cv_results, function(x) x$rmse))
mean_mape <- mean(sapply(cv_results, function(x) x$mape))
sd_rmse <- sd(sapply(cv_results, function(x) x$rmse))
sd_mape <- sd(sapply(cv_results, function(x) x$mape))

cat("Mean RMSE:", mean_rmse, "\n")
cat("Mean MAPE:", mean_mape, "%\n")
cat("SD RMSE:", sd_rmse, "\n")
cat("SD MAPE:", sd_mape, "%\n")
cat("Number of folds:", length(cv_results), "\n")
```

The linear model was retained as a benchmark. The final thesis model used the
nonlinear `nlsLM` specification because the nonlinear approach provided a more
flexible representation of media saturation and carryover effects.
